---
title: "Operating model demonstration"
author: "Charles T T Edwards"
date: "`r format(Sys.Date(), '%d-%b-%Y')`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    toc: no
vignette: >
  %\VignetteIndexEntry{Operating model demonstration}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = 'fig/om-examples-', fig.width = 6, tidy = TRUE, tidy.opts = list(blank = TRUE, width.cutoff = 95), message = FALSE, warning = FALSE, collapse = TRUE, comment = "#>")
```

```{r, message = TRUE}
library(om)
```
```{r}
library(dplyr)
library(ggplot2)
```

## Setup

Setup operating model class object, assuming 3 samples from the input life-history distribution and a default projection period of 200 years. 
 
```{r}
om_object <- om(ages = 0:12, samples = 3L, time = 201)
```
```{r, echo = FALSE}
#print(om_object@seeds)
om_object@seeds <- c(2032574L, 4149433L, 9193660L)
```

Set-up list of parameter distributions required for population dynamics function:
```{r}
om_object <- om_object %>% load_pars(list(
    's' = distribution(pars = c(2.2, 0.2), density = "logitnormal", name = "adult survivorship"),
    'b' = distribution(pars = c(0.0, 0.2), density = "lognormal", name = "female fecundity"),
    'm' = distribution(value = 4, name = "age at maturity"),
    'l' = distribution(value = 0.7, name = "immature survivorship multiplier")
))
```

Each parameter is stored as a `distribution` class object and assigned to the `om_object@pars` slot. A simple plotting function can be used to quickly visualise the parameter input distributions. For example:
```{r}
plot(om_object@pars$s)
```

The intrinsic growth rate is calculated automatically when parameters are assigned and stored in `om_object@pars$r`. There exists a `lambda()` function for extracting the intrinsic growth on either the natural or logarithmic scale. For example:
```{r}
plot(lambda(om_object))
```

The intrinsic growth can be assumed to be the maximum, or an alternative can be supplied. The default behaviour assigns `om_object@pars$r` to `om_object@pars$rmax` but truncates the distribution at zero:
```{r}
om_object <- om_object %>% load_rmax()

plot(om_object@pars$rmax)
```

To estimate the reference points, we are also required to load the observation and selectivity paramter distributions:
```{r}
om_object <- om_object %>% update_pars(list(
    'o' = distribution(value = 1, name = "age at observation"),
    'v' = distribution(value = 4, name = "age at selectivity")
))
```

By default the 1+ carrying capacity is set to one:
```{r}
om_object@pars$K
```

For the projections, uncertainty is required:
```{r}
# load uncertainty
om_object <- om_object %>% load_cvs(list(
    'survivorship' = 0.05,
    'birth'        = 0.01,
    'observation'  = 0.00,
    'mortality'    = 0.00)
)
```

## Fixed shape

Given the input parameter distributions, the shape parameter determines the depeletion at MNPL. This can be illustrated using a fixed shape value. 

```{r}
# assume shape
shape(om_object) <- 1

# deterministic ref. points
om_object0 <- rp(om_object, stochastic = FALSE, time = 200L)

targets(om_object0)
```

The `spf()` function can be useful for plotting the surplus production functions being assumed by the model:
```{r}
dfr0 <- spf(om_object0, harvest_rate = seq(0.00, max(targets(om_object0)$harvest_rate$value) * 2, length = 101))
dfr1 <-  data.frame(sample = as.character(1:om_object@samples), harvest_rate = om_object0@targets$harvest_rate, captures = om_object0@targets$captures, depletion = om_object0@targets$depletion)
ggplot() + 
  geom_line(data = dfr0, aes(x = harvest_rate, y = captures, col = sample)) +
  geom_vline(data = dfr1, aes(xintercept = harvest_rate, col = sample), linetype = "dashed") +
  geom_point(data = dfr1, aes(x = harvest_rate, y = captures, col = sample))
ggplot() + 
  geom_line(data = dfr0, aes(x = depletion, y = captures, col = sample)) +
  geom_vline(data = dfr1, aes(xintercept = depletion, col = sample), linetype = "dashed") +
  geom_point(data = dfr1, aes(x = depletion, y = captures, col = sample))
```

When performing a stochastic projection, uncertainty is a consequence of the life-history parameter distributions only. By default, projections take place assuming a harvest rate equal the harvest rate at MNPL:
```{r}
om_object0@harvest_rate
```

```{r}
# population dynamics
om_object0 <- pdyn(om_object0, stochastic = FALSE, initial_depletion = 0.1)
dynplot(om_object0)
```

Diagnostics are available per sample:
```{r}
diagnostics(om_object0)
```

## Estimated shape

The shape can be estimated per life-history sample using the `shape` function, in which case the desired target depletion must be specified:
```{r}
# estimate shape for assumed d_mnpl assumption
om_object1 <- shape(om_object, depletion = 0.7, stochastic = FALSE, time = 200L)
shape(om_object1)
```

As before reference points can be estimated and the surplus production function(s) plotted:
```{r}
# deterministic ref. points
om_object1 <- rp(om_object1)
targets(om_object1)

dfr0 <- spf(om_object1, harvest_rate = seq(0.00, max(targets(om_object1)$harvest_rate$value) * 2, length = 101))
dfr1 <-  data.frame(sample = as.character(1:om_object@samples), harvest_rate = om_object1@targets$harvest_rate, captures = om_object1@targets$captures, depletion = om_object1@targets$depletion)
ggplot() + 
  geom_line(data = dfr0, aes(x = harvest_rate, y = captures, col = sample)) +
  geom_vline(data = dfr1, aes(xintercept = harvest_rate, col = sample), linetype = "dashed") +
  geom_point(data = dfr1, aes(x = harvest_rate, y = captures, col = sample))
ggplot() + 
  geom_line(data = dfr0, aes(x = depletion, y = captures, col = sample)) +
  geom_vline(data = dfr1, aes(xintercept = depletion, col = sample), linetype = "dashed") +
  geom_point(data = dfr1, aes(x = depletion, y = captures, col = sample))
```

Projection confirms that the population will converge on the depletion at MNPL specified during estimate of the shape:
```{r}
om_object1 <- pdyn(om_object1, stochastic = FALSE, initial_depletion = 0.3, use_rmax = TRUE)
dynplot(om_object1)

diagnostics(om_object1)
```
